---
title: "How to use the `mybcart` package"
author: Bruna Wundervald
date: March, 2019
bibliography: bibliography.bib
output: pdf_document
editor_options: 
  chunk_output_type: console
header-includes: \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE, 
                      fig.pos="H")
```

The `mybcart` package is a very slow implementation of the Bayesian 
CART model, made mostly with functions of the `tidyverse` collection
of packages. The package can be installed from GitHub with:

```{r, eval = FALSE}
devtools::install_github("brunaw/my_BART", subdir = "package/mybcart")
```


After that, a list of the available functions can be obtained with:

```{r}
library(mybcart)
ls("package:mybcart")
```


## Brief description of the functions

There are 11 functions in the package (besides the `%>%` and the
Friedman data), that can be described as: 

  - `bcart`: a function that receives a formula, the data and the 
  number of iterations to build the model, and depends on: 
    - `lk_ratio_grow`: Calculates the likelihood ratio of a grow step;
    - `lk_ratio_prune`: Calculates the likelihood ratio of a prune step;
    - `structure_ratio_grow`: Calculates the tree structure ratio of a grow step;
    - `structure_ratio_prune`: Calculates the tree structure ratio of a prune step;
    - `transition_ratio_grow`: Calculates the transition ratio of a grow step;
    - `transition_ratio_prune`: Calculates the transition ratio of a prune step;
    - `ratio_grow`: Combines the tree ratios of a grow step, to calculate the final 
    acceptance value for the MCMC; 
    - `ratio_prune`: Combines the tree ratios of a prune step, to calculate the final 
    acceptance value for the MCMC; 
    - `p_rule`: A function that selects the splitting rule in a grow step, 
    given a variable uniformly selected. 
    
  - `predict_bcart`: A function that performs the prediction for a 
  `bcart` model, by receiveing the model and the new data. 
  
More details about each function can be found in their documentation 
inside the package, 
as well as the details for the ratios and the model should be found
in the maths file. The `bcart` function will return a list containing: 

  - `sigma_2`: The posterior values for $\sigma^2_{y}$;
  -  `errors`: The sum of the squared residuals calculated for
  each iteration of the model;
  - `final_tree`: The final tree;
  - `ratios`: The calculated ratios in each step; 
  - `samp_unif`: The uniform values sampled to used in the MCMC; 
  - `nodes`: The nodes, which were grown or pruned, used in each 
  iteration of the model. 
  - `action_taken`: The action (prune, grow or nothing) taken
  in each iteration. 
  - `results`: The order in which the growing happened for the
  final tree; 
  - `mu`: The values for $\mathbf{\mu}$ for the final tree, 
  used in the prediction function. 
  - `model_formula`: The model formula, used in the prediction
  function. 
  
## The Friedman data

The package also contains some data to be used with the model.
The data was simulated using the model equation proposed in 
(@Friedman1991). We simulated a response variable $Y$ and its 
relationship to a matrix of predictors $\mathbf{X}$ as

\begin{equation}
y_i = 10 sin(\pi x_{i1} x_{i2}) + 20 (x_{i3} -
0.5)^{2} + 10 x_{i4} + 5 x_{i5} +
 \epsilon_i, \thinspace
\epsilon_i \stackrel{iid}\sim N(0, \sigma^2),
\end{equation}

where $\mathbf{X} \in [0, 1]$, meaning that the predictors
were randomly drawn from a standard Uniform distribution.  

## How to use it 

The following code show how to actually use the functions, starting by
gathering the data, running the model and checking it, and 
using the prediction function. 

```{r}
library(tidyverse)

# Gathering the data and separating in train and test 
data <- mybcart::friedman_data %>% 
  mutate(set = ifelse(runif(nrow(.)) > 0.8, "test", "train"))

train <- data %>% filter(set == "train") %>% select(-set)

# Running the model
bcart_model <- mybcart::bcart(formula = y ~ ., data = train, iter = 200)
str(bcart_model)
```

```{r, fig.cap="Chain for $\\sigma^{2}$ and density of sum of squared errors for a B-CART model run with 200 iterations", fig.align='center', fig.width=7, fig.height=3.5}
# Checking some results of the model
library(patchwork)
p1 <- data.frame(sig = bcart_model$sigma_2,
           ind = 1:201) %>% 
  ggplot(aes(ind, sig)) +
  geom_line(size = 0.30) +
  labs(x = "Iterations", 
       y = expression("Posterior values of"~sigma[y]^{2})) +
  theme_bw()


p2 <- data.frame(err = bcart_model$errors) %>% 
  ggplot(aes(err)) +
  geom_density(fill = "sienna1") +
  labs(x = "Sum of squared errors", 
       y = "Density") + 
  theme_bw()

p1 + p2 + plot_layout(nrow = 1)
```

```{r}
# Using the prediction function 
test <- data %>% filter(set == "test") %>% select(-set)
pred <- predict_bcart(model = bcart_model, newdata = test)

pred %>% 
  group_by(node) %>% 
  summarise(ssr = sum((y - prediction)^2),
            mean_of_y = mean(y), 
            mu_prediction = mean(prediction))
```
  
## Regex dictionary 

Some regular expression were used in the pruning part of the model. They can 
be described as:

  - `'( right| left)$'`: detects the last "right" or "left";
  - `'X[0-9][^X[0-9]]*$'`: detects the last X followed by a number;
  - `'(?<='something'\\s).*'`: detects everything after the "something";
  - `'( X[0-9] right| X[0-9] left)$'`: detects the 
  last X followed by a number and a "left" or "right";
  